---
title: "Two_group_brain_samples"
author: "Foong Lian Chee", edited by "Shenfeng Qiu"
date: "2/28/2022" # DO NOT CHANGE THIS CODE, shinny bubble plot works!!
output: html_document
editor_options: 
  chunk_output_type: console
---
# v2 has statistical comparison within each cluster, still developing...
# v3 has additional features of ploting heatmaps, v4 refines v3
# Analyzing scRNA-seq Data

## Initialize relevant libraries
```{r}
# [You have to change] set this following work directory to your file location
setwd("D:/Bioinformatics Projects/LindseyAanalysis/single_cell/two_brain_groups/")
library(Seurat)
library(dplyr)
library(cowplot)
library(ggplot2)
library(rstatix)

library(ComplexHeatmap)
library(circlize)
#if (!requireNamespace("BiocManager", quietly = TRUE))
 #   install.packages("BiocManager")

#BiocManager::install("SingleCellExperiment")
library(SingleCellExperiment)

```

## Import Group samples 
```{r}
# You might need to change the location "slim_counts/", to direct to your data folder
for (file in c("m09",
               "m10",
               "m11",
               "m12")){
  seurat_data <- Read10X(data.dir = paste0("D:/Bioinformatics Projects/LindseyAanalysis/single_cell/slim_counts/", file))
  seurat_obj <- CreateSeuratObject(counts = seurat_data, 
                                   min.features = 100, 
                                   project = file)
  assign(file, seurat_obj)
  
               }
m09 <- AddMetaData(object = m09, metadata = "m09", col.name = "sample.id")
m10 <- AddMetaData(object = m10, metadata = "m10", col.name = "sample.id") 
m11 <- AddMetaData(object = m11, metadata = "m11", col.name = "sample.id")
m12 <- AddMetaData(object = m12, metadata = "m12", col.name = "sample.id")
m09 <- AddMetaData(object = m09, metadata = "WT", col.name = "group.id") 
m10 <- AddMetaData(object = m10, metadata = "WT", col.name = "group.id")
m11 <- AddMetaData(object = m11, metadata = "KO", col.name = "group.id") 
m12 <- AddMetaData(object = m12, metadata = "KO", col.name = "group.id")

gallitano <- merge(m09, y = c(m10,m11,m12), 
                   add.cell.ids = c("m09","m10","m11", "m12"), 
                   project = "two_groups")
gallitano[["percent.mt"]] <- PercentageFeatureSet(gallitano, pattern = "^mt-") 

rm(m09, m10, m11, m12)
rm(seurat_obj, seurat_data)
head(gallitano)
tail(gallitano)
table(gallitano@meta.data$sample.id)
# m09   m10   m11   m12 
# 5194 15772  9932  4357 
table(gallitano@meta.data$group.id)
# KO    WT 
#14289 20966 
```

# Quality Control
## Plot Data

```{r}
# For all the png function in the following, try to replace the url address with your output folder
# you can use the "Ctrl + F" the search bar, to replace all the relevant folder address
png(filename = "./v3_output/QC_beforefiltering.png",
    width = 800, height = 1000)
VlnPlot(gallitano, 
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        group.by = "sample.id",
        ncol = 1) + ggtitle("Quality assessment before filtering")
dev.off()

library(patchwork)
png(file = "./v3_output/QC_FeatureScatter.png", width = 1000, height = 500) #skip this line to output to console, or after png file is generated and dev.off, then run the following lines to output to console
plot1 <- FeatureScatter(gallitano, feature1 = "nCount_RNA", feature2 = "percent.mt") 
line.data <- data.frame(yintercept = c(1500, 7500), Lines = c("lower", "upper"))
line2.data <- data.frame(xintercept = c(500), Lines = c("lower"))

plot2 <- FeatureScatter(gallitano, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + geom_hline(aes(yintercept = yintercept, linetype = Lines), line.data) + geom_vline(aes(xintercept = xintercept, linetype = Lines), line2.data)

plot1 + plot2 +  plot_layout(guides = 'collect') + plot_annotation(title = "Quality assessment before filtering")
dev.off()

```

# Joint filtering effects:

```{r}
# joint filtering effect
# Visualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIs
metadata <- gallitano@meta.data
png(file = "./v3_output/QC_jointfilteringeffect.png", width = 600, height = 500) #skip this line to output to console, or after png file is generated and dev.off, then run the following lines to output to console
metadata %>% 
  	ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 500) +
  	geom_hline(yintercept = c(1500, 7500))
dev.off()

rm(metadata, line.data,line2.data,plot1,plot2)
```

## Selecting Data Subset
# I follow the original script to filter based on these criteria:
nFeature_RNA < 7500& nFeature_RNA>1500 & nCount_RNA > 500 
# and then, I added the filtering criteria of "percent.mt < 20"
# After subseting, I also remove the genes with zero count in more than 100 cells
```{r}
gallitano <- subset(gallitano, nFeature_RNA < 7500 & 
                      nFeature_RNA>1500 & nCount_RNA > 500 & 
                      percent.mt < 20)

# Output a logical vector for every gene on whether the more than zero counts per cell
# Extract counts
counts <- GetAssayData(object = gallitano, slot = "counts")
# Output a logical vector for every gene on whether there more than zero counts per cell
nonzero <- counts > 0
# Sums all TRUE values and returns TRUE if more than 100 TRUE values per gene
keep_genes <- Matrix::rowSums(nonzero) >= 100
# Only keeping those genes expressed in more than 100 cells
filtered_counts <- counts[keep_genes, ]
# Reassign to filtered Seurat object
gallitano <- CreateSeuratObject(filtered_counts, meta.data = gallitano@meta.data)

rm(filtered_counts,nonzero, counts, keep_genes, file)
```

## Reaccess quality metrics
```{r}
dim(gallitano)
# 14377 genes x 11424 cells (or UMIs)

table(gallitano@meta.data$sample.id)
# m09  m10  m11  m12 
# 1966 3633 3896 1929

png(file = "./v3_output/QC_afterfiltering.png", width = 600, height = 1200)
VlnPlot(gallitano, 
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), 
        ncol = 1) + ggtitle("Quality assessment after filtering")
dev.off()

png(file = "./v3_output/QC_FeatureScatter_after.png", width = 1000, height = 500)
plot1 <- FeatureScatter(gallitano, feature1 = "nCount_RNA", feature2 = "percent.mt") 
line.data <- data.frame(yintercept = c(1500, 7500), Lines = c("lower", "upper"))
line2.data <- data.frame(xintercept = c(500), Lines = c("lower"))

plot2 <- FeatureScatter(gallitano, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + geom_hline(aes(yintercept = yintercept, linetype = Lines), line.data) + geom_vline(aes(xintercept = xintercept, linetype = Lines), line2.data)

plot1 + plot2 +  plot_layout(guides = 'collect') + plot_annotation(title = "Quality assessment after filtering")
dev.off()

metadata <- gallitano@meta.data
png(file = "./v3_output/QC_jointfilteringeffect_after.png", width = 600, height = 500)
metadata %>% 
  	ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 500) +
  	geom_hline(yintercept = c(1500, 7500))
dev.off()

rm(metadata, line.data,line2.data, plot1, plot2)
```

## Apply sctransform normalization
This steps can replace NormalizeData(), ScaleData(), and FindVariableFeatures(). The results of sctransform are stored in the "SCT" assay. It is assumed to reveals sharper biological distinctions compared to the standard Seurat workflow. 

Since it is a normalization step, we have to do separately for two different samples (that's why split seurat objects here), then only integrate the sample expression data in the next step, integration analysis, to remove batch effect.

```{r}
options(future.globals.maxSize = 9000 * 1024^2) #100 gb

# Split seurat object by condition to perform cell cycle scoring and SCT on all samples
split_seurat <- SplitObject(gallitano, split.by = "sample.id")

split_seurat <- split_seurat[c("m09","m10", "m11", "m12")]

for (i in 1:length(split_seurat)) {
  split_seurat[[i]] <- NormalizeData(split_seurat[[i]], verbose = TRUE)
    split_seurat[[i]] <- SCTransform(split_seurat[[i]],method = "glmGamPoi", vars.to.regress = c("percent.mt"))
}

rm(gallitano)

```

## Integration 

```{r}
# Select the most variable features to use for integration
integ_features <- SelectIntegrationFeatures(object.list = split_seurat, 
                                            nfeatures = 3000) 
# Prepare the SCT list object for integration
split_seurat <- PrepSCTIntegration(object.list = split_seurat, 
                                   anchor.features = integ_features)
# Find best buddies - can take a while to run
integ_anchors <- FindIntegrationAnchors(object.list = split_seurat, 
                                        normalization.method = "SCT", 
                                        anchor.features = integ_features)
# Integrate across conditions
seurat_integrated <- IntegrateData(anchorset = integ_anchors, 
                                   normalization.method = "SCT")

rm(integ_anchors)
rm(integ_features)
rm(split_seurat)
```
## Clustering
```{r}
# Run the standard workflow for visualization and clustering
seurat_integrated <- RunPCA(object = seurat_integrated, npcs = 40, verbose = F)
ElbowPlot(seurat_integrated, ndims = 40)
PCAPlot(seurat_integrated)  
seurat_integrated <- RunUMAP(seurat_integrated, 
                             dims = 1:40,
                             reduction = "pca")
seurat_integrated <- RunTSNE(seurat_integrated, 
                             reduction = "pca")
       # Determine the K-nearest neighbor graph
seurat_integrated <- FindNeighbors(object = seurat_integrated, 
                                   dims = 1:40)

# Determine the clusters for various resolutions                                
seurat_integrated <- FindClusters(object = seurat_integrated,
                                  resolution = c(0.05, 0.5))
#setwd("D:/Bioinformatics Projects/LindseyAanalysis/single_cell/two_brain_groups")
#saveRDS(seurat_integrated, "./v3_output/Seurat_integrated_twobraingroups.RDS")
```

## Clustering quality control
This step gives us some idea about how is the distribution of the number of genes, number of UMIs, and percentage of mitochondrial genes in each cluster. 
Normally, we expect to see similar distribution of no. of genes (nFeature_RNA) and no. of UMIs (nCount_RNA). 

As for the percent.mt (percentage of mitochondrial genes per cell), it can be a reference to check if those high intensity clusters might be having poor quality cells (if so, we can try to remove in the next step or adjust the metrics in the previous filtering step) or it might be due to the differences biologically

```{r}
# Determine metrics to plot present in seurat_integrated@meta.data
metrics <-  c("nFeature_RNA", "nCount_RNA", "percent.mt")

png(file = "./v3_output/QC_Featureplot_after.png", width = 1000, height = 800)
FeaturePlot(seurat_integrated, 
            reduction = "umap", 
            features = metrics,
            split.by = "sample.id",
            pt.size = 0.4,
            min.cutoff = 'q10',
            label = TRUE)
dev.off()
```

# additional plotting
```{r}
Idents(object = seurat_integrated) <- seurat_integrated$integrated_snn_res.0.05
# Idents(object = seurat_integrated) <- seurat_integrated$sample.id
# Idents(object = seurat_integrated) <- seurat_integrated$group.id
png(file = "./v3_output/DimPlot_tsne.png", width = 1000, height = 800)
DimPlot(seurat_integrated, reduction = "tsne", label = TRUE, pt.size = 1, label.size = 6)
dev.off()

png(file = "./v3_output/DimPlot_umap.png", width = 1000, height = 800)
DimPlot(seurat_integrated, reduction = "umap", label = TRUE, pt.size = 1, label.size = 6)
dev.off()

png(file = "./v3_output/PCAPlot.png", width = 500, height = 400)
PCAPlot(seurat_integrated)
dev.off()

# do feature plots for selected genes
select_genes = c('Trem2','Snap25','Epas1', 'Egr1', 'Npas4', 'Calcrl') 
png(file = "./v3_output/FeaturePlot_select_genes_tsne.png", width = 1000, height = 800)
FeaturePlot(seurat_integrated, features =select_genes, cols =c('gray','red'), reduction ='tsne', pt.size = 0.7,ncol = 3, label = T)
dev.off()
png(file = "./v3_output/FeaturePlot_select_genes_umap.png", width = 1000, height = 800)
FeaturePlot(seurat_integrated, features =select_genes, cols =c('gray','red'), reduction = 'umap',pt.size = 0.7,ncol=3, label = T)
dev.off()
```
# do heatmaps
```{r}
select_genes2                 <- c('Snap25','Gad1','Gad2', 'Slc32a1', 'Slc17a7', 'Lamp5', 'Aqp4', 'Gfap', "Ndnf", 'Sncg', 'Vip', 'Trem2','Sst','Pvalb', 'Cux2', 'Rorb', 'Fezf2', 'Foxp2', 'Npas4', 'Mbp', 'Cldn5', 'Ctss', 'C1qa')
cellRanks                <- seurat_integrated@meta.data[order(seurat_integrated@meta.data$integrated_snn_res.0.05),] # check cellRanks structure!
PartialMatrix            <- seurat_integrated@assays$integrated@scale.data[which(rownames(seurat_integrated@assays$integrated@scale.data) %in% select_genes2), rownames(cellRanks)]
cellRanks$col            <- rainbow(max(as.numeric(cellRanks$integrated_snn_res.0.05))+1)[as.numeric(cellRanks$integrated_snn_res.0.5)+1]
#
ha_column <- HeatmapAnnotation(
  df  = data.frame(
    ClusterID = as.numeric(cellRanks$integrated_snn_res.0.05)
  ),
  col = list(
    ClusterID = colorRamp2(unique(as.numeric(cellRanks$integrated_snn_res.0.05)), 
                           rainbow(max(as.numeric(cellRanks$integrated_snn_res.0.05))))
  )
)

png(file = "./v3_output/select_geneset_heatmap.png", width = 1000, height = 400)
ht1 = Heatmap(PartialMatrix, name = "Scaled\nUMI", column_title = "Allen SMARTseq dataset genes", 
              top_annotation = ha_column, show_column_names=FALSE, cluster_columns = FALSE,
              row_names_gp = gpar(fontsize = 20))
# do not use TRUE for clustering, too many cells cannot see clearly
draw(ht1)
# ht1
dev.off()

# heatmap, without gene name cluster, adopt listed gene sequences
png(file = "./v3_output/heatmap_noCluster.png", width = 1000, height = 400)
DoHeatmap(seurat_integrated, features = select_genes2)
dev.off()

png(file = "./v3_output/heatmap_groupby_groupID.png", width = 1000, height = 400)
DoHeatmap(seurat_integrated, features = select_genes2, group.by = 'group.id' )
dev.off()
#DoHeatmap(seurat_integrated, cells = 1:100, features = select_genes2, group.by = 'group.id' )
png(file = "./v3_output/heatmap_groupby_seurat_clusters.png", width = 1000, height = 400)
DoHeatmap(seurat_integrated, features = select_genes2, group.by = 'seurat_clusters' )
dev.off()
```
# additiontal plots, adjusting format
```{r}
Idents(object = seurat_integrated) <- seurat_integrated$integrated_snn_res.0.5
png(file = "./v3_output/DimPlot with finer res.png", width = 600, height = 400)
DimPlot(seurat_integrated, label = T, pt.size = 1, label.size = 6)
dev.off()

png(file = "./v3_output/Vlnplot4selectedFeatures.png", width = 1000, height = 300)
plot <- VlnPlot(seurat_integrated, features = c("Gad1","Trem2"), combine = FALSE, fill.by = c("feature","ident")) 
plot <-  lapply(
  X = plot,
  FUN = function(p) p + aes(color= seurat_integrated$integrated_snn_res.0.5)
)
CombinePlots(plots = plot, legend = 'none')
dev.off()

```
# additional custermized heatmap plot
```{r}
Idents(object = seurat_integrated) <- seurat_integrated$integrated_snn_res.0.05
# or for finer clusters:
# Idents(object = seurat_integrated) <- seurat_integrated$integrated_snn_res.0.5
png(file = "./v3_output/heatmap_customize.png", width = 1200, height = 800)
DoHeatmap(seurat_integrated, features = select_genes2, disp.min = -1,
          slot = 'scale.data', 
          group.colors = rainbow(9)) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red",midpoint = 1)
dev.off()

```
# trial DEGs and GSEA analysis for a cluster. Note: the following block can run, but it takes long time and i am not sure it means anythig...so let's skip this block.
```{r}
# Idents(object = seurat_integrated) <- seurat_integrated$integrated_snn_res.0.05

# PG.markers.7 <- FindMarkers(seurat_integrated, ident.1 = "0", min.pct = 0.20) # we will adopt the variable name .7, but to represent cluster 0.
#PG.markers.7$gene <- rownames(PG.markers.7)
#library(biomaRt)

#gene_id <- getBM(attributes = c("ensembl_gene_id","external_gene_name","entrezgene_id"), mart = useDataset("mmusculus_gene_ensembl", useMart("ensembl"))) # takes longer time
#PG.markers.7 <- merge(PG.markers.7, gene_id[,c(2,3)], by.x = "gene", by.y = "external_gene_name")
#PG.markers.7.filt <- PG.markers.7
#PG.markers.7.filt <- PG.markers.7.filt[!duplicated(PG.markers.7.filt$entrezgene_id),]
#PG.markers.7.filt <- PG.markers.7[which(PG.markers.7$p_val_adj < 0.05),]
#genelist <- PG.markers.7$avg_log2FC
#names(genelist) <-  PG.markers.7$entrezgene_id
#genelist <- sort(genelist, decreasing = TRUE)

#library(fgsea)
#library(msigdbr)
#library(data.table)
#library(tidyverse)

#m_df <- msigdbr(species = "Mus musculus", category = "C2") %>% dplyr::select(gs_name, entrez_gene)
#m_df_fgsea <- split(x = m_df$entrez_gene, f = m_df$gs_name)

#fgseaRes <- fgsea(pathways = m_df_fgsea, 
                  #stats    = genelist,
                  #eps      = 0.05,
                  #minSize  = 15,
                  #maxSize  = 800) # takes longer ~40min for cluster0 genes
#selected <- fgseaRes[c(979, 329, 443, 445, 907),]
#selected <- selected$pathway

#png(file = "./v3_output/fgsea table.png", width = 800, height = 500)
#plotGseaTable(m_df_fgsea[selected], genelist, fgseaRes, 
             # gseaParam=0.5)
#dev.off()
```

# Find all markers in two samples for cell type identification

```{r}
Idents(seurat_integrated)<- "seurat_clusters"
Qiu.markers <-  FindAllMarkers(seurat_integrated,
                                    only.pos=TRUE, min.pct=0.25, logfc.threshold=0.25)
write.csv(Qiu.markers,"./v3_output/Qiu_all_markers_20220226.csv")
```

## Identifying cell type
## Option 1: SingleR package with built-in reference
I use a collection of mouse bulk RNA-seq data sets obtained from celldex package (Benayoun et al. 2019). A variety of cell types are available, mostly from blood but also covering several other tissues. This identifies marker genes from the reference and uses them to compute assignment scores (based on the Spearman correlation across markers) for each cell in the test dataset against each label in the reference. The label with the highest score is the assigned to the test cell, possibly with further fine-tuning to resolve closely related labels.

This reference consists of a collection of mouse bulk RNA-seq data sets downloaded from the gene expression omnibus (Benayoun et al. 2019). A variety of cell types are available, again mostly from blood but also covering several other tissues.


```{r}
#BiocManager::install("SingleR")
#BiocManager::install("celldex")
library(SingleR)
library(celldex)

ref <- MouseRNAseqData()
table(Idents(object = seurat_integrated))

hpca.se <- SingleR(test = seurat_integrated@assays$RNA@data, ref = ref,
                    labels = ref$label.main)

png(filename = "./v3_output/plotScoreHeatmap_m10.png",
    width = 600, height = 500)
p1 <- plotScoreHeatmap(hpca.se)
dev.off()

png(filename = "./v3_output//plotDeltaDistribution_m10.png",
    width = 1000, height = 1400)
plotDeltaDistribution(hpca.se, ncol = 3)
dev.off() # if plot does not complete, dev.off again to get null device 1 report

all.markers <- metadata(hpca.se)$de.genes
seurat_integrated$labels_hpca <- hpca.se$labels
# Extract identity and sample information from seurat object to determine the number of cells per cluster per sample
n_cells <- FetchData(seurat_integrated, 
                     vars = c("integrated_snn_res.0.5",
                              "labels_hpca")) %>%
  dplyr::count(integrated_snn_res.0.5,
               labels_hpca) %>%
  tidyr::spread(integrated_snn_res.0.5, n)

# View(n_cells)
write.csv(n_cells, "./v3_output/n_cells_mouseRNAseqData_twobraingroups.csv")
dev.off()

## comparing with the unsupervised clustering
tab <- table(cluster= seurat_integrated$integrated_snn_res.0.5, label= hpca.se$labels) 

png(filename = "./v3_output/pheatmap_SingleR_m10.png",
    width = 600, height = 500)
p3 <- pheatmap::pheatmap(log10(tab+10))

dev.off()
```


## Option 2: manual annotation
# refer to the seurat_integrated original script
Plotting Astrocyte Markers
```{r}
png(filename = "./v3_output/astrocyte_marker_score.png", width = 1500, height = 400)
VlnPlot(object = seurat_integrated, features =c("Aqp4","Prdx6","Pla2g7"))
dev.off()
```

Plotting Microglia Markers
```{r}
png(filename = "./v3_output/microglia_marker_score.png", width = 1500, height = 400)
VlnPlot(object = seurat_integrated, features =c("C1qc","Ly86", "Tmem119"))
dev.off()
```

Plotting Endothelial Markers
```{r}
png(filename = "./v3_output/endothelial_marker_score.png", width = 1500, height = 400)
VlnPlot(object = seurat_integrated, features =c("Ly6a",  "Ly6c1",  "Pltp"))
dev.off()
```
Plotting Oligodendrocyte Markers
```{r}
png(filename = "./v3_output/oligoDC_marker_score.png", width = 1500, height = 400)
VlnPlot(object = seurat_integrated, features =c("Mag",  "Mbp",  "Cldn11"))
dev.off()
```

Plotting Glutamatergic Neuron Markers
```{r}
png(filename = "./v3_output/Glutamatergic_Neuron_marker_score.png", width = 1500, height = 400)
VlnPlot(object = seurat_integrated, features =c("Nrgn", "Sv2b",  "Snap25"))
dev.off()
```
Plotting Gabaergic Neuron Markers
```{r}
png(filename = "./v3_output/Gabaergic_Neuron_marker_score.png", width = 1500, height = 400)
VlnPlot(object = seurat_integrated, features =c("Gad1",  "Gad2",  "Rab3b" ))
dev.off()
```
Plotting Oligodendrocyte Precursor Markers
```{r}
png(filename = "./v3_output/OligoDC_Precursor_marker_score.png", width = 1500, height = 400)
VlnPlot(object = seurat_integrated, features =c("Pdgfra",  "Cspg4",  "Gm2a"))
dev.off()
```

## Rename cluster based on the SingleR results and manual annotation

```{r}
seurat_integrated <- RenameIdents(object = seurat_integrated, 
                         "0" = "Neurons(Glutamatergic)",
                         "1" = "Astrocytes",
                         "2" = "Endothelial cells",
                         "3" = "Neurons(Glutamatergic)",
                         "4" = "Microglia",
                         "5" = "Neurons(Glutamatergic)",
                         "6" = "Oligodendrocytes",
                         "7" = "Neurons(Glutamatergic)",
                         "8" = "Neurons(Glutamatergic)",
                         "9" = "OPC_precursor",
                         "10" = "Neurons(Gabaergic)",
                         "11" = "Neurons(Glutamatergic)",
                         "12" = "Neurons(Gabaergic)",
                         "13" = "Neurons(Glutamatergic)",
                         "14" = "Neurons(Glutamatergic)",
                         "15" = "Endothelial cells",
                         "16" = "Endothelial cells",
                         "17" = "Neurons(Glutamatergic)",
                         "18" = "Astrocytes",
                         "19" = "Neurons(Glutamatergic)",
                         "20"= "Astrocytes",
                         "21"= "Neurons(Glutamatergic)",
                         "22"= "Oligodendrocytes",
                         "23"= "Endothelial cells",
                         "24"= "Endothelial cells"                                                  )
seurat_integrated$celltype_LC <- seurat_integrated@active.ident

png(filename = "./v3_output/UMAP_res0.5.png", 
    width = 500, height = 400)
DimPlot(seurat_integrated,
                reduction = "umap",
                group.by = "integrated_snn_res.0.5",
                label = TRUE,
)
dev.off()

png(filename = "./v3_output/UMAP_res0.5_samples.png", 
    width = 1200, height = 500)
DimPlot(seurat_integrated,
                reduction = "umap",
                group.by = "integrated_snn_res.0.5",
                split.by = "sample.id",
                label = TRUE,
)
dev.off()

Idents(object = seurat_integrated) <- "celltype_LC"

plot <- DimPlot(seurat_integrated,
                reduction = "umap",
                group.by = "celltype_LC",
                label = FALSE,
)
plot$data$seurat_clusters <- seurat_integrated@meta.data$integrated_snn_res.0.5
all(rownames(plot$data) == rownames(seurat_integrated@meta.data))
png(filename = "./v3_output/UMAP_res0.5_celltype_LC.png", 
    width = 600, height = 400)
LabelClusters(plot, id = "seurat_clusters")
dev.off()

Idents(object = seurat_integrated) <- "celltype_LC"
plot <- DimPlot(seurat_integrated,
                reduction = "umap",
                group.by = "celltype_LC",
                split.by = "sample.id",
                label = FALSE,
)
plot$data$seurat_clusters <- seurat_integrated@meta.data$integrated_snn_res.0.5
all(rownames(plot$data) == rownames(seurat_integrated@meta.data))
png(filename = "./v3_output/UMAP_res0.5_celltype_LC_sample.png", 
    width = 1200, height = 500)
LabelClusters(plot, id = "seurat_clusters")
dev.off()

saveRDS(seurat_integrated, "./v3_output/seurat_integrated_twobraingroups.RDS")
```


# Differential expression Analysis
```{r}
Idents(seurat_integrated) <- "group.id"
groups <- unique(Idents(seurat_integrated)) # here clusters is a misnomer? -actually ABCD four groups

pairwise <- combn(groups, 2)

results <- list()
# one comparison group WT and KO that takes 25 min
for(i in 1:ncol(pairwise)) {
  markers <- FindMarkers(seurat_integrated, ident.1 = pairwise[1, i], ident.2 = pairwise[2, i], logfc.threshold = 0)
  comparisons <- pairwise[, i]
  markers$comparison <- paste(comparisons[1], comparisons[2], sep = '_')
  results[[i]] <- markers
}

results.df <- do.call(rbind, results)
saveRDS(results.df, "./v3_output/between_group_comp.rds")
```


```{r}
Idents(seurat_integrated)<- "group.id"

clusters <- levels(seurat_integrated$celltype_LC)

groups <- unique(Idents(seurat_integrated))
pairwise <- combn(groups, 2)

full_results<-data.frame()
for (k in 1:length(clusters)){
  results <- list()
  seurat_integrated.subset <- subset(seurat_integrated, subset= celltype_LC == clusters[k])
  for(i in 1:ncol(pairwise)) {
    markers <- FindMarkers(seurat_integrated.subset, ident.1 = pairwise[1, i], ident.2 = pairwise[2, i], logfc.threshold = 0)
    comparisons <- pairwise[, i]
    markers$comparison <- paste(comparisons[1], comparisons[2], sep = '_')
    markers$genes<- rownames(markers)
    results[[i]] <- markers
  }
  results.df <- do.call(rbind, results)
  results.df$Cluster<- clusters[k]
  full_results<-rbind(full_results,results.df)
}
  
full_results
saveRDS(full_results, "./v3_output/full_results.rds")

write.csv(full_results, "./v3_output/full_results_marker_comp_by_clusters.csv")
```
## SQ: not sure if we need to do this, manual annotation already has great results
## Option 3: manual annotation for 25 clusters
Refer to Tasic et al, Nature 2018, marker list
# https://github.com/AllenInstitute/tasic2018analysis/blob/master/RNA-seq%20Analysis/markers.R


```{r}
Inh.markers <- list(c("Gad2","Slc32a1","Prox1","Adarb2","Nfix","Nfib","Cacna2d1",
                  "Cxcl14","Tnfaip8l3","Cplx3","Lamp5","Cd34","Pax6","Krt73",
                  "Scrg1","Egln3","Ndnf","Tmem182","Ntn1","Pde11a","Pdlim5",
                  "Lsp1","Slc35d3","Nkx2-1","Serpinf1","Col14a1","Vip","Sncg",
                  "Crabp1","Slc10a4","Cldn10","Bhlhe22","Crispld2","Slc17a8",
                  "Cyb5r2","Nr1h4","Wnt7b","Prss12","Igfbp6","Calb2","Grpr",
                  "Pthlh","Elfn1","Rspo1","Slc18a3","Lmo1","Rspo4","Sostdc1",
                  "Chat","Cbln4","Gsx2","Gpc3","Mab21l1","C1ql1","Itih5","Mybpc1",
                  "Myl1","Lhx6","Sox6","Sst","Chodl","Calb1","Cbln4","Etv1","Edn1",
                  "Kl","Il1rapl2","Myh8","Ptprk","Chrna2","Myh13","Ptgdr","Crhr2",
                  "Hpse","Igsf9","C1ql3","Tacstd2","Th","Col6a1","Nts","Tac1","Pvalb",
                  "Gabrg1","Il7","Bche","Prdm8","Syt2","Ostn","Pdlim3","C1ql1",
                  "Gpr149","Vipr2","Meis2","Adamts19","Cpa6","Lgr6"))
Inh.comb.markers <- c("Reln","Cnr1","Nr2f2","Cck","Npy","Crh","Tac2")

Ex.markers <- list(unique(c("Slc17a7","Rtn4rl2","Slc30a3","Cux2","Stard8","Otof","Rrad",
                            "Penk","Agmat",
                      "Emx2","S100a3","Macc1","Rorb","Scnn1a","Whrn","Endou","Col26a1",
                      "Rspo1","Fezf2","Hsd11b1","Batf3","Arhgap25","Colq","Pld5","Olfr78",
                      "Tcap","Fgf17","Wfdc18","Wfdc17","Aldh1a7","Tgfb1","Ctsc","Rxfp2",
                      "Prss35","Rgs12","Osr1","Oprk1","Cd52","Col23a1","Col18a1","Car1",
                      "Car3","Fam84b","Chrna6","Chrnb3","Fn1","Tac1","Lce3c","Erg",
                      "Cdc42ep5","Bmp5","Pvalb","Depdc7","Stac","C1ql2","Ptgfr","Slco2a1",
                      "Pappa2","Dppa1","Npsr1","Htr2c","Hpgd","Nxph3","Sla2","Tshz2",
                      "Rapgef3","Slc17a8","Trh","Nxph2","Foxp2","Col12a1","Syt6","Col5a1",
                      "Gpr139","Ly6d","Sla","Cpa6","Ppp1r18","Faim3","Ctxn3","Nxph4",
                      "Cplx3","Ctgf","Col8a1","Mup5","Ngf","Fam150a","F2r","Serpinb11","Fbxl7",
                      "P2ry12","Crh","Kynu","Hsd17b2","Mup3","Tlcd1","Lhx5","Trp73","Cpa6",
                      "Gkn1","Col18a1","Lce3c","Erg","Bmp5","Stac","C1ql2","Slco2a1","Lrrc9",
                      "Trhr","Myzap","Krt80","H60b","Fam150a","Clic5","Kcnj5","Olfr110",
                      "Olfr111")))
Ex.comb.markers <- c("Reln","Cdh13","Cpne7","Alcam","Rprm","Marcksl1")
                            

Global.markers <- list(c("Fez1","Phyhipl","Aplp1","Gnao1","Caly","Snap25","Atp1a3","Camk2b",
                    "Syt1","Gabrg2","Fabp3","Stmn2","Kif5c","Slc32a1","Gad2","Dlx1","Dlx5",
                    "Dlx2","Dlx6os1","Slc6a1","Sox2","Slc17a7","Nrn1","Neurod2","Sv2b","Satb2",
                    "Tbr1","Vsig2","Cmtm5","Kcnj10","S100a16","S100a13","S1pr1","Gja1","Gjb6",
                    "Aqp4","Lcat","Acsbg1","Olig1","Sox10","Neu4","Sapcd2","Gpr17","Plp1",
                    "Cldn11","Mag","Mog","Nkx6-2","Enpp6","9630013A20Rik","Brca1",
                    "Mog","Opalin","Gjb1","Hapln2","Cyba","Ctsh","Ifitm3","Sparc",
                    "S100a11","Dcn","Col1a1","Pltp","Vtn","Slc6a13","Spp1","Slc13a3",
                    "Col15a1","Slc47a1","Tgtp2","Ifi47","Esam","Slco1a4","Slc38a5",
                    "Cldn5","H2-Q7","Slc38a11","Art3","Ace2","Acta2","Myh11","Pln",
                    "Gja5","Kcnj8","Atp13a5","Aoc3","Ctss","C1qb","C1qc","C1qa","Cbr2",
                    "F13a1","Pf4","Mrc1","Siglech","Selplg"))

seurat_integrated <- AddModuleScore(object = seurat_integrated, 
                           features = Ex.markers, 
                           name = "Excitatory_marker_score")
png(filename = "./v3_output/UMAP_Excitatory_marker_score.png", width = 700, height = 700)
FeaturePlot(object = seurat_integrated, features = "Excitatory_marker_score1", label = T) + 
  ggtitle("FeaturePlot_Excitatory marker score (refer to Tasic et al., 2018, Nature)")
dev.off()

seurat_integrated <- AddModuleScore(object = seurat_integrated, 
                           features = Inh.markers, 
                           name = "Inhibitory_marker_score")
png(filename = "./v3_output/UMAP_Inhibitory_marker_score.png", width = 700, height = 700)
FeaturePlot(object = seurat_integrated, features = "Inhibitory_marker_score1", label = T) + 
  ggtitle("FeaturePlot_Inhibitory marker score (refer to Tasic et al., 2018, Nature)")
dev.off()

seurat_integrated <- AddModuleScore(object = seurat_integrated, 
                           features = Global.markers, 
                           name = "Global_marker_score")
png(filename = "./v3_output/UMAP_Global_marker_score.png", width = 700, height = 700)
FeaturePlot(object = seurat_integrated, features = "Global_marker_score1", label = T) + 
  ggtitle("FeaturePlot_Global marker score (refer to Tasic et al., 2018, Nature)")
dev.off()

saveRDS(seurat_integrated, "./v3_output/seurat_integrated_twobraingroups.RDS")
```

# additional plots
```{r}
Qiu.ab<- subset(seurat_integrated, subset = group.id %in% c("WT", "KO"))
ab_markers<- subset(full_results, subset = comparison == "WT_KO") 
ab_markers<- subset(ab_markers, subset = p_val_adj < 0.005)
ab_markers<- ab_markers[order(-abs(ab_markers$avg_log2FC)),]
write.csv(ab_markers, "./v3_output/sorted_ab_marker_all_clusters.csv")
##
num_markers<- as.data.frame(table(ab_markers$Cluster))
num_markers<- num_markers[order(num_markers$Freq),]
png(filename = "./v3_output/num_DEG_by_cluster_in_AB.png", width = 800, height = 800)
p<-ggplot(data=num_markers, aes(x=Var1, y=Freq, fill=Var1)) +
  geom_bar(stat="identity")+
  theme_minimal()+
  coord_flip()+
  labs(title="Number of Differentially Expressed Genes by Cluster in AB", y="Number of Differentially Expressed Genes", x="Cell Type")+
  theme(legend.position="none")
p
dev.off()

##
Idents(Qiu.ab)<-'celltype_LC'
features<- unique(ab_markers$genes[1:11])
png(filename = "./v3_output/DotPlot_top_markers_AB.png", width = 800, height = 800)
DotPlot(Qiu.ab, features = features) + RotatedAxis()
dev.off()

##
Idents(Qiu.ab)<-'celltype_LC'
#current.cluster.ids <- c("Glutamatergic","Astrocyte","Endothelial","Microglia","Oligodendrocyte Precursor","Gabaergic","Pericyte","Oligodendrocyte","Endothelial+Pericyte")
#new.cluster.ids <- c("Gl_N","A","EC","M","O PC","Gb_N","P","O","EC+P")
#Qiu.ab@active.ident <- plyr::mapvalues(x = Qiu.ab@active.ident, from = current.cluster.ids, to = new.cluster.ids)
#Qiu.ab@meta.data$cell.type.abbr <- Qiu.ab@active.ident
png(filename = "./v3_output/Heatmap_top100_markers_AB.png", width = 2000, height = 1000)
DoHeatmap(subset(Qiu.ab, downsample = 100), features = unique(ab_markers$genes)[1:100], size = 3)+
      theme(axis.text.y = element_text(size = 5), text=element_text(size=20))
dev.off()

```


```{r}

# library(monocle3)
# library(SeuratWrappers) 

#data <- GetAssayData(seurat_integrated)[VariableFeatures(seurat_integrated),]
#pd <- data.frame(seurat_integrated@meta.data)
#fData <- data.frame(gene_short_name = rownames(data), row.names = row.names(data))
#cds <- new_cell_data_set(expression_data = data,
                       # cell_metadata = pd,
                       # gene_metadata = fData)
#cds <- preprocess_cds(cds, num_dim = 12,method = "PCA")
#plot_pc_variance_explained(cds)
#cds <- reduce_dimension(cds,
                        #preprocess_method = "PCA",
                        #reduction_method= "tSNE",
                        #verbose = T)
#cds <- cluster_cells(cds = cds , reduction_method = "tSNE", verbose = TRUE,
                     #resolution = NULL)
#p1 <- plot_cells(cds, reduction_method = "tSNE")
##compare with seurat data, res = 0.5
#cds@clusters$tSNE$clusters <- seurat_integrated$integrated_snn_res.0.5
#p2 <- plot_cells(cds, reduction_method = "tSNE")
#p1 + p2

```


# visualization using ShinnyCell


```{r}
library(ShinyCell)

setwd("D:/Bioinformatics Projects/LindseyAanalysis/single_cell/two_brain_groups/v3_output")
# getExampleData()    # Download example dataset (~200 MB) no need to run this if you run it before, with the .rds file in working directory
seurat_integrated = readRDS("seurat_integrated_twobraingroups.RDS")
scConf = createConfig(seurat_integrated)
makeShinyApp(seurat_integrated, scConf, gene.mapping = TRUE,
             shiny.title = "m9 m10 m11 m12 two brain groups") 

# rm(list = ls())
# rm(list = setdiff(ls(), "plot"))
setwd("D:/Bioinformatics Projects/LindseyAanalysis/single_cell/two_brain_groups/v3_output/shinyApp") # verify change of dir in console using getwd()

shiny::runApp() # after you run this App, another R will be open, maximize it, then select and plot and test the graphic output. While this R window is open, the > ready in console will NOT be ready. You can also select output to browser. In order for the browser to work, you need to keep this new R window open. If you close the R window, the brower will gray out. 
```


